---
title: "Bird Recognition"
output: html_notebook
---

## Load the necessary packages

```{r}
library(bioacoustics)
library(tuneR)
library(seewave)
library(dplyr)
library(tools)
library(randomForest)
library(stringr)
library(keras)
```

## Load the files into a data frame. Separate out the metadata from the filename into columns. Render some spectrograms

```{r}
files = list.files("wav_files_playback", "*.wav", full.names=TRUE)
files_without_extension = basename(file_path_sans_ext(files))
wavs = setNames(
  lapply(
    files,
    read_audio
  ),
  files_without_extension
)
metadata = data.frame(str_split_fixed(files_without_extension, "_", 3))
metadata = cbind(metadata, files_without_extension)
colnames(metadata) = c("birdid", "calltype", "idnumber", "filename")
head(metadata)
oscillo(wavs[[1]])
birds = unique(metadata$birdid)
filtered = filter(metadata, birdid == birds[1])
for (i in 1:nrow(filtered)) {
  audio = wavs[[filtered$filename[i]]]
  spectro(audio, main=filtered$filename[i])
}
```

Each calltype seems to have a unique spectrogram.

## Split dataset into train / test. The test dataset will contain one row for each combination of birdid / calltype

```{r}
test = metadata %>% 
          group_by(birdid, calltype) %>% 
          filter(row_number()==1)
test
train = metadata %>% 
          group_by(birdid, calltype) %>% 
          filter(row_number()!=1)
train
```

## Train a random forest classifier based on Mel-frequency Cepstral Coefficients (MFCC)

```{r}
mels = sapply(wavs, melfcc, numcep = 20) # Calculate all MFCCs
head(mels[[1]]) # Result is a matrix of 20 coefficients across nrow time frames
image(mels[[1]])

trainMels = mels[train$filename] # Select the MFCCs corresponding to the training dataset
trainM = do.call(rbind, trainMels) # melfcc gives a matrix - rbind to cast from 3D to 2D across the whole training dataset
y = rep(train$birdid, lapply(trainMels, function(x) dim(x)[1])) # Create birdid labels for each MFCC
rownames(trainM) = y
head(trainM)

set.seed(42)
rf = randomForest(trainM, y, importance = FALSE, proximity = FALSE, replace = TRUE, ntree = 4000, mtry = 4)
rf

testMels = mels[test$filename]
testM = do.call(rbind, testMels)
y = rep(test$birdid, lapply(testMels, function(x) dim(x)[1]))
rownames(testM) = y
head(testM)

# Let's make predictions with our classifier on a test set
table(y, predict(rf, testM, type = "response"))

# To look at the predictions 
predict(rf, testM, type = "prob")

predictions = character(nrow(test))
for (i in 1:nrow(test)) {
  mel = testMels[[i]]
  prediction = names(head(sort(colMeans(predict(rf, mel, type="prob")), decreasing = TRUE)))[1]
  predictions[i] = prediction
}
testWithPredictions = data.frame(test, predictions)
testWithPredictions
correct_predictions = nrow(filter(testWithPredictions, birdid == predictions))
accuracy_pct = round(correct_predictions / nrow(test) * 100, 2)
print(paste(correct_predictions, "/", nrow(test), "wavs in the test dataset correctly identifed. Accuracy: ", accuracy_pct, "%"))
```

## Deep learning classification with the R interface to Keras based on MFCC

```{r}

```

