---
title: "Time sync check"
output: html_notebook
---

## Load libraries

```{r message=FALSE, warning=FALSE}
# install.packages(c("bioacoustics", "tuneR", "seewave", "dplyr", "tidyverse", "lubridate", "tools", "ggplot2", "ggpmisc", "plotly", "stringr", "monitoR", "pbapply", "foreach", "doParallel", "R.utils"), type="binary")
library(bioacoustics)
library(tuneR)
library(seewave)
library(dplyr)
library(tidyverse)
library(lubridate)
library(tools)
library(ggplot2)
library(ggpmisc)
library(plotly)
library(stringr)
library(foreach)
library(doParallel)
library(R.utils)
library(readxl)
library(pbapply)
ymd_hms = function(dt) {
  lubridate::ymd_hms(dt, tz="Pacific/Auckland")
}
dmy_hms = function(dt) {
  lubridate::dmy_hms(dt, tz="Pacific/Auckland")
}
options(stringsAsFactors = FALSE)
```

## Load RFID log files

```{r}
files = list.files("data/2019 RFID", pattern = "*.TXT$", recursive = TRUE, full.names=TRUE)
csvs = lapply(files, function(path) {
  csv = read.csv(path, header=FALSE, col.names = c("year", "month", "day", "hour","minute_sec", "id", "ext", "int", "msg"))
  csv$site = strsplit(strsplit(path, "/")[[1]][4], "_")[[1]][1]
  csv
})
rfid = do.call(rbind, csvs)
rfid$datetime = paste0(rfid$year, "-", rfid$month, "-", rfid$day, " ", rfid$hour, ":", rfid$minute_sec)
rfid$datetime = ymd_hms(rfid$datetime)
ggplot(rfid, aes(datetime)) +
  geom_freqpoly(binwidth = 60 * 60)
```
### Reference spectrograms for playback files
```{r}
html_tag_audio <- function(file, type = c("wav")) {
  type <- match.arg(type)
  htmltools::tags$h3(
    basename(file),
    htmltools::tags$audio(
      controls = "",
      htmltools::tags$source(
        src = file,
        type = glue::glue("audio/{type}", type = type)
      )
    )
  )
}
files = Sys.glob("./clips/PB*.wav")
PB_wavs = lapply(files, function(f) {
  wav = read_audio(f)
  wav_length = round(length(wav) / wav@samp.rate, 2)
  spectro(wav, main=paste0(basename(f), " ", wav_length, "s"), osc=TRUE)
  print(html_tag_audio(f))
  data.frame(filename=basename(f), wav_length)
})
PB_wavs = do.call(rbind, PB_wavs)
```

### Build a reference dataframe showing which wave files correspond to which times

```{r}
read_BAR_metadata = function(files) {
  library(pbapply)
  df = pblapply(files, function(f) {
    bits = strsplit(basename(f), "_")[[1]]
    start = ymd_hms(paste(bits[2], bits[3]))
    site = strsplit(f, "/")[[1]][4]
    try({
      wav_header = readWave(f, header=TRUE)
      duration = wav_header$samples / wav_header$sample.rate
      end = start + duration
      return(data.frame(filename = f, base_filename = basename(f), site = site, start = start, end = end, samples = wav_header$samples, sr = wav_header$sample.rate, duration=duration, interval = interval(start, end)))
    })
    NULL
  })
  df = do.call(rbind, df)
  df
}
if (file.exists("misc/BAR_timespans.csv")) {
  df = read.csv("misc/BAR_timespans.csv")
  df$start = ymd_hms(df$start)
  df$end = ymd_hms(df$end)
  df$interval = interval(df$start, df$end)
} else {
  files = list.files("data/2019 BAR recordings", pattern = "*.wav$", recursive = TRUE, full.names=TRUE)
  df = read_BAR_metadata(files)
  write.csv(df, "misc/BAR_timespans.csv", row.names = FALSE)
}
df[c("site", "interval")]
```


### Select a few playback entries to check

```{r}
PB_to_check = "PB0023.wav"
site_to_check = "NNK30"
filtered_rfid = filter(rfid, msg == PB_to_check & site == site_to_check)
filtered_df = filter(df, site == site_to_check)
filtered_rfid$corrected_datetime = filtered_rfid$datetime
filtered_rfid$match = lapply(1:nrow(filtered_rfid), function(i) {
  dt = filtered_rfid$corrected_datetime[i]
  site = filtered_rfid$site[i]
  indices = which(dt %within% filtered_df$interval)
  if (length(indices) == 0) indices = NA
  indices
})
print(paste("There are", nrow(filtered_rfid), PB_to_check, "playbacks, of which", sum(!is.na(filtered_rfid$match)), "have recordings"))
# A few randomly
datetimes_to_check = sort(sample(filtered_rfid$corrected_datetime[!is.na(filtered_rfid$match)], 5))
# Ones from the 13th
#datetimes_to_check = filtered_rfid$corrected_datetime[!is.na(filtered_rfid$match) & filtered_rfid$datetime > ymd_hms("2019-11-20 00:00:00") & filtered_rfid$datetime < ymd_hms("2019-11-20 23:59:59")]
#datetimes_to_check = filtered_rfid$corrected_datetime[!is.na(filtered_rfid$match) & filtered_rfid$corrected_datetime %within% interval(ymd_hms("2019-12-13 16:45:00"), ymd_hms("2019-12-13 17:13:00"))]
print(datetimes_to_check)
```

```{r}
#dt = ymd_hms("2019-12-13 16:45:04")
setWavPlayer("play") # OS dependent
checked_clips = lapply(datetimes_to_check, function(dt) {
  match = filtered_rfid$match[filtered_rfid$corrected_datetime == dt][[1]]
  f = filtered_df$filename[match]
  clip_start = time_length(dt - filtered_df$start[match], unit = "seconds")
  clip_end = clip_start + 1
  print(paste(f, dt, clip_start, clip_end))
  wav = read_audio(f, from = clip_start, to = clip_end)
  spectro(wav, main=dt, osc=TRUE)
  play(wav)
  wav
})
```

## Write out a csv of approx positions of PB events per wave file

```{r}
if (file.exists("misc/approx_offsets.csv")) {
  approx_offsets = read.csv("misc/approx_offsets.csv")
} else {
  filtered_rfid = rfid[str_extract(rfid$msg, "^.{2}") == "PB",]
  approx_offsets = lapply(1:nrow(df), function(i) {
    matched_events = filtered_rfid[filtered_rfid$site == df$site[i] & filtered_rfid$datetime %within% df$interval[i],]
    if (nrow(matched_events) == 0) {
      return(NULL)
    }
    if (nrow(matched_events) > 10) {
      matched_events = sample_n(matched_events, 10)
    }
    offsets = time_length(matched_events$datetime - df$start[i])
    data.frame(filename = df$filename[i], offset = offsets, site = matched_events$site, pb = matched_events$msg, datetime = matched_events$datetime)
  })
  approx_offsets = do.call(rbind, approx_offsets)
  write.csv(approx_offsets, "misc/approx_offsets.csv", row.names = FALSE)
}
```

## Compare annotations with nearest RFID log

```{r}
annotations = lapply(list.files("misc", "*.txt", full.names=TRUE), function(f) {
  if (str_detect(f, "ZOOM")) {
    return(NULL)
  } else if (str_detect(f, "selections")) {
    x = read.table(f, sep="\t", header=TRUE)
    x$datetime = ymd_hms(paste(x$Begin.Date, x$Begin.Clock.Time))
    x = x[c("Begin.Time..s.", "End.Time..s.", "Individual.ID", "Nest.ID", "datetime")]
  } else {
    x = read.table(f, col.names = c("Begin.Time..s.", "End.Time..s.", "Individual.ID"))[c(TRUE, FALSE),]
    x$Begin.Time..s. = as.numeric(x$Begin.Time..s.)
    x$End.Time..s. = as.numeric(x$Begin.Time..s.)
    x$Nest.ID = strsplit(basename(f), "_")[[1]][1]
    x$datetime = df$start[df$base_filename == str_replace(basename(f), ".txt", ".wav")] + x$Begin.Time..s.
  }
  x$filename = paste0(str_replace(basename(f), ".Band.Limited.Energy.Detector.selections.txt|.txt", ""), ".wav")
  x
})
annotations = do.call(rbind, annotations)
annotations$offset = rep(NA, nrow(annotations))
annotations$Individual.ID[annotations$Individual.ID == "PB0009" | annotations$Individual.ID == "PB009"] = "PB0009zp"
annotations$Nest.ID[annotations$Nest.ID == "NKN31"] = "NNK31"
for (i in 1:nrow(annotations)) {
  if (substring(annotations$Individual.ID[i], 0, 2) == "PB") {
    filtered_rfid = filter(rfid, annotations$Nest.ID[i] == site & msg == paste0(annotations$Individual.ID[i], ".wav"))
    matched_rec_start = df$start[df$site == annotations$Nest.ID[i] & annotations$datetime[i] %within% df$interval]
    diffs = time_length(filtered_rfid$datetime - annotations$datetime[i])
    min_i = which.min(abs(diffs))
    offset = diffs[min_i]
    if (length(offset) == 1) {
      annotations$offset[i] = offset
    }
  }
}

approx_offsets = read_excel("misc/approx_offsets.xlsx")
names(approx_offsets) = c("filename", "guide_start", NA, NA, "Begin.Time..s.", "Nest.ID", "Individual.ID", "datetime", NA)
annotations = within(annotations, rm(End.Time..s.))
approx_offsets$offset = approx_offsets$guide_start - approx_offsets$Begin.Time..s.
approx_offsets$filename = basename(approx_offsets$filename)
approx_offsets$Individual.ID = str_replace(approx_offsets$Individual.ID, ".wav", "")
annotations = rbind(annotations, approx_offsets[names(annotations)])

annotations = na.omit(annotations)

filtered_annotations = filter(annotations, offset > -60 & Nest.ID == "NNS1")
ggplot(data = filtered_annotations, aes(x=Begin.Time..s., y=offset, color=substr(basename(filename), start = 6, stop = 20))) +
    geom_point() +
    geom_smooth(method = "lm") +
    stat_poly_eq(formula = y ~ x,
                   aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                   parse = TRUE) +
    labs(color='file datetime')
filtered_annotations = filter(annotations, Nest.ID == "NNBB12")
ggplot(data = filtered_annotations, aes(x=Begin.Time..s., y=offset)) +
    geom_point() +
    geom_smooth(method = "lm") +
    stat_poly_eq(formula = y ~ x,
                   aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                   parse = TRUE)
ggplot(data = annotations, aes(x=Begin.Time..s., y=offset, color=Nest.ID)) +
    geom_point() +
    geom_smooth(method = "lm") +
    stat_poly_eq(formula = y ~ x,
                   aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                   parse = TRUE)
```

Something is wrong with the RFID chip clock - it's drifting significantly. More data points will help - let's autodetect PB in the audio files

```{r}
filtered_annotations = annotations[annotations$Individual.ID == "PB0023" & annotations$Nest.ID == "NNK30",]
print(filtered_annotations[1,])
library(monitoR)
PB23 = makeCorTemplate("clips/PB0023.wav", frq.lim = c(3,10), name="PB0023")
PB24 = makeCorTemplate("clips/PB0024.wav", frq.lim = c(3,10), name="PB0024")
ctemps = combineCorTemplates(PB23, PB24)
wav = read_audio(filtered_annotations$Begin.Path[1], from = 300, to = 800)
savewav(wav, filename="test.wav")
#spectro(wav, osc=TRUE)
#play(wav)
cscores <- corMatch("test.wav", ctemps, parallel = TRUE, show.prog = TRUE)
cdetects <- findPeaks(cscores)
getDetections(cdetects)
```
monitoR seems too unreliable for this task, unfortunately


The reason for the drift is SD card write speed. BAR logfiles indicate the true time packed, so if we read the BAR logfiles we can determine the true sampling rate, and correct it. In some cases, it'll be just under 44.1KHz.

```{r}
files = list.files("data/2019 BAR recordings", pattern = "logfile.*txt$", recursive = TRUE, full.names=TRUE)
sites = sapply(strsplit(files, "/"), "[[", 4)
df$true_length = rep(NA, nrow(df))
for (site in unique(sites)) {
  files_for_site = files[site == sites]
  lines = unlist(lapply(files_for_site, readLines))
  rec = str_subset(lines, "Recording")
  dt = dmy_hms(str_extract(rec, "^.{19}"))
  rec = rec[order(dt)]
  dt = sort(dt)
  for (i in 1:(length(rec) - 1)) {
    if (str_detect(rec[i], "started") && str_detect(rec[i + 1], "stopped")) {
      filtered_df = df[df$site == site,]
      diffs = time_length(filtered_df$start - dt[i])
      min_i = which.min(abs(diffs))
      min_i = which(df$filename == filtered_df$filename[min_i])
      df$true_length[min_i] = time_length(dt[i+1] - dt[i])
    }
  }
}
ggplot(df[df$site != "KAKA Aviary_20191101_20191113_BAR",], aes(x=start, y=true_length - duration, color = site)) +
  geom_point() + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

## Write out waves with corrected sampling rates

```{r eval=FALSE, include=FALSE}
registerDoParallel(detectCores() - 1)
filtered_df = df[!is.na(df$true_length) & df$true_length > 100 & df$site %in% unique(rfid$site),]
foreach(i = 1:nrow(filtered_df), .packages = "tuneR")  %dopar% {
  output_filename = str_replace(filtered_df$filename[i], "2019 BAR recordings", "2019 BAR recordings - fixed SR")
  if (!file.exists(output_filename)) {
    wav = readWave(filtered_df$filename[i])
    wav@samp.rate = round(length(wav) / filtered_df$true_length[i])
    mkdirs(dirname(output_filename))
    writeWave(wav, filename = output_filename)
  }
}
```

The problem with that though, is that sampling rates in wave files are integers - there's a loss of precision there. It should be possible to multiply a desired second offset by the true sampling rate and slice a wav file that way.

```{r}
df$true_sr = df$samples / df$true_length
df$true_sr[!is.na(df$true_sr) & df$true_sr > 44000] = 44100
write.csv(df, "misc/BAR_timespans.csv", row.names = FALSE)
ggplot(df[df$duration > 100,], aes(x=true_sr)) +
  geom_histogram(binwidth=5)
```
```{r}
approx_offsets$corrected_offset = sapply(1:nrow(approx_offsets), function(i) {
  approx_offsets$offset[i] * 44100 / df$true_sr[approx_offsets$filename[i] == df$filename]
})
write.csv(approx_offsets, "misc/approx_offsets.csv", row.names = FALSE)
```


```{r}
filtered_offsets = approx_offsets[approx_offsets$site == "NNS1" & approx_offsets$filename %in% df$filename[!is.na(df$true_sr)],]
for (i in sample(1:nrow(filtered_offsets), 10)) {
  f = filtered_offsets$filename[i]
  true_sr_for_f = df$true_sr[df$filename == f]
  clip_start = round(filtered_offsets$offset[i] * true_sr_for_f)
  clip_end = round(clip_start + true_sr_for_f)
  print(paste(f, true_sr_for_f, clip_start, clip_end))
  wav = readWave(filename = f, from = clip_start, to = clip_end, units="samples")
  wav@samp.rate = true_sr_for_f
  spectro(wav, main=basename(f), osc=TRUE)
  play(wav)
}
```

```{r}
annotations$corrected_begin = sapply(1:nrow(annotations), function(i) {
  true_sr = df$true_sr[str_detect(df$base_filename, fixed(annotations$filename[i]))]
  annotations$Begin.Time..s.[i] * 44100 / true_sr
})
annotations$offset = NA
for (i in 1:nrow(annotations)) {
  if (substring(annotations$Individual.ID[i], 0, 2) == "PB") {
    filtered_rfid = filter(rfid, annotations$Nest.ID[i] == site & msg == paste0(annotations$Individual.ID[i], ".wav"))
    matched_rec_start = df$start[str_detect(df$base_filename, fixed(annotations$filename[i]))]
    rfid_offset = time_length(filtered_rfid$datetime - matched_rec_start)
    diffs = time_length(annotations$corrected_begin[i] - rfid_offset)
    min_i = which.min(abs(diffs))
    offset = diffs[min_i]
    if (length(offset) == 1) {
      annotations$offset[i] = offset
    }
  }
}
filtered_annotations = filter(annotations, Nest.ID == "NNS1" & abs(offset) < 60)
ggplot(data = filtered_annotations, aes(x=corrected_begin, y=offset, color=substr(basename(filename), start = 6, stop = 20))) +
    geom_point() +
    geom_smooth(method = "lm") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    labs(color='file datetime') +
    stat_poly_eq(formula = y ~ x,
                   aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                   parse = TRUE) +
    ggtitle("Drift per file at NNS1")

ggplot(data = filtered_annotations, aes(x=datetime, y=offset)) +
    geom_point() +
    geom_smooth(method = "lm") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    labs(color='file datetime') +
    stat_poly_eq(formula = y ~ x,
                   aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                   parse = TRUE) + 
    ggtitle("Drift per day at NNS1")

ggplot(data = annotations, aes(x=corrected_begin, y=offset, color=Nest.ID)) +
    geom_point() +
    geom_smooth(method = "lm") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10))

filtered_annotations = filter(annotations, Nest.ID == "NNBB12")
ggplot(data = filtered_annotations, aes(x=datetime, y=offset)) +
    geom_point() +
    geom_smooth(method = "lm") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    labs(color='file datetime') +
    stat_poly_eq(formula = y ~ x,
                   aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                   parse = TRUE) + 
    ggtitle("Drift per day at NNBB12")
```

```{r}
f = "NNS1_20191203_064040_SunriseToSunset [-39.0986 176.8061]"
offset = median(annotations$offset[annotations$filename == f], na.rm=TRUE) # offset for this file
matched_df = df[str_detect(df$base_filename, fixed(f)),]
matched_event_offsets = time_length(rfid$datetime[rfid$msg == "PB0023.wav" & rfid$datetime %within% matched_df$interval & rfid$site == matched_df$site] - matched_df$start)
for (begin_time in matched_event_offsets) {
  sr = matched_df$true_sr
  clip_start = round((begin_time + offset) * sr)
  clip_end = round(clip_start + (sr * 2))
  print(paste(matched_df$filename, matched_df$true_sr, clip_start, clip_end))
  wav = readWave(filename = matched_df$filename, from = clip_start, to = clip_end, units="samples")
  wav@samp.rate = true_sr_for_f
  spectro(wav, main=begin_time, osc=TRUE)
  #play(wav)
}
```

Not bad.

## Let's rename these RFID codes with something a bit more readable. We'll need to maintain a way of converting between.


```{r}
unique_rfids = unique(rfid$msg[str_extract(rfid$msg, "^.{2}") == "RF"])
#human_readable_rfids = paste0("bird", 1:length(unique_rfids))
human_readable_rfids = c("alpha", "bravo", "charlie", "delta", "echo", "foxtrot", "golf", "hotel", "india", "juliet", "kilo", "lima", "mike", "november")
human_readable_rfids = setNames(human_readable_rfids, unique_rfids)
unique_rfids = setNames(unique_rfids, human_readable_rfids)

processed_rfid = rfid
mask = str_extract(processed_rfid$msg, "^.{2}") == "RF"
processed_rfid$msg[mask] = human_readable_rfids[processed_rfid$msg[mask]]
ggplot(processed_rfid[mask,], aes(x = datetime, color = msg)) +
  geom_freqpoly(binwidth = 60 * 60 * 24) + 
  ggtitle("RFID detections per day")
```

## Try detect entry / exit events by looking at light gate changes

```{r}
if (file.exists("misc/processed_rfid.csv")) {
  processed_rfid2 = read.csv("misc/processed_rfid.csv")
  processed_rfid2$datetime = ymd_hms(processed_rfid2$datetime)
} else {
  time_threshold = 2
  LG_threshold = 1
  library(pbapply)
  processed_rfid2 = pblapply(1:sum(mask), function(i) {
    dt = processed_rfid$datetime[mask][i]
    this_site = processed_rfid$site[mask][i]
    bird = processed_rfid$msg[mask][i]
    LG = filter(processed_rfid, site == this_site & abs(datetime - dt) < time_threshold & msg == "Light Gate Change")
    if (nrow(LG) > 2) {
      event_type = NULL
      ext_g = mean(diff(LG$ext))
      int_g = mean(diff(LG$int))
      if (ext_g < 0 & int_g > 0) {
        event_type = "entry"
      } else if (int_g < 0 & ext_g > 0) {
        event_type = "exit"
      }
      if (length(event_type) == 1) {
        print(paste(i, bird, event_type, this_site, dt))
        return(data.frame(i = i, bird = bird, event_type = event_type, ext_g = ext_g, int_g = int_g, site = this_site, datetime = dt))
      }
    }
  })
  processed_rfid2 = do.call(rbind, processed_rfid)
  write.csv(processed_rfid2, "misc/processed_rfid.csv", row.names=FALSE)
}
ggplot(processed_rfid2, aes(x = datetime, y = bird, color = event_type)) + geom_point()
ggplot(filter(processed_rfid2, datetime > ymd("2019-12-28") & datetime < ymd("2019-12-29")), aes(x = datetime, y = bird, color = event_type)) + geom_point()
```
Seems a bit too unreliable - the light gates might be too sensitive. Looks like we'll have to just slice around RFID passes.

## Build a dataframe of datetime intervals to slice - intelligently overlapping for same bird id, and removing intervals where there might be confusion (a playback, or another bird)

```{r}
time_threshold = 5
events_to_ignore = c("Power-on Reset", "Low-voltage Detect Reset", "Light Gate Change", 
  "Watchdog(COP) Reset", "Playback Error! Restarting", "Software Reset"
)

library(pbapply)
dts_to_slice = pblapply(1:sum(mask), function(i) {
    dt = processed_rfid$datetime[mask][i]
    this_site = processed_rfid$site[mask][i]
    bird = processed_rfid$msg[mask][i]
    this_id = processed_rfid$id[mask][i]
    
    start = dt - time_threshold
    end = dt + time_threshold
    nearby_events = filter(processed_rfid, id != this_id & site == this_site & abs(time_length(datetime - dt)) < time_threshold & !(msg %in% events_to_ignore))
    other_birds = human_readable_rfids[!human_readable_rfids == bird]
    if (any(other_birds %in% nearby_events$msg)) {
      print(paste(i, "another bird within threshold"))
      return(NULL)
    }
    if (any(str_extract(nearby_events$msg, "^.{2}") == "PB")) {
      print(paste(i, "playback within threshold"))
      return(NULL)
    }
    data.frame(i = i, site = this_site, bird = bird, start = start, end = end, interval = interval(start, end))
})
dts_to_slice = do.call(rbind, dts_to_slice)
write.csv(dts_to_slice, "misc/dts_to_slice.csv", row.names = FALSE)

dts_to_slice_deduplicated = dts_to_slice %>% group_by(site, bird) %>% arrange(start) %>% 
                      mutate(indx = c(0, cumsum(as.numeric(lead(start)) >
                              cummax(as.numeric(end)))[-n()])) %>%
                      group_by(site, bird, indx) %>%
                      summarise(start = min(start), end = max(end)) %>%
                      select(-indx)
dts_to_slice_deduplicated$duration = time_length(dts_to_slice_deduplicated$end - dts_to_slice_deduplicated$start)
write.csv(dts_to_slice_deduplicated, "misc/dts_to_slice_deduplicated.csv", row.names = FALSE)
ggplot(dts_to_slice_deduplicated, aes(duration)) + geom_histogram(binwidth = 1)
```

## Put median offset per file in df
## For each interval around a RFID pass, output a wave clip, corrected for offset and SR

```{r}
df$offset = sapply(df$base_filename, function(f) {
  median(annotations$offset[paste0(annotations$filename) == f], na.rm = TRUE)
})
pblapply(1:nrow(dts_to_slice_deduplicated), function(i) {
  start = dts_to_slice_deduplicated$start[i]
  duration = dts_to_slice_deduplicated$duration[i]
  bird = dts_to_slice_deduplicated$bird[i]
  site = dts_to_slice_deduplicated$site[i]
  formatted_date = format(start, "%Y-%m-%d_%H%M%S_NZDT")
  output_filename = paste0("data/2019 BAR recordings - individual bird clips/", bird, "/", site, "_", formatted_date, ".wav")
  if (file.exists(output_filename)) {
    return(NULL)
  }
  matched_df = df[df$site == site & start %within% df$interval & !is.na(df$offset) & !is.na(df$true_sr),]
  if (nrow(matched_df) == 1) {
    sr = matched_df$true_sr
    clip_start_s = time_length(start - matched_df$start) + matched_df$offset
    clip_start_samples = round(clip_start_s * sr)
    clip_end_samples = round(clip_start_samples + (sr * duration))
    #print(paste(i, bird, site, start, duration, matched_df$filename, matched_df$true_sr, clip_start_samples, clip_end_samples))
    try({
      wav = readWave(filename = matched_df$filename, from = clip_start_samples, to = clip_end_samples, units="samples")
      wav@samp.rate = sr
      if (sr != 44100) {
        wav = resamp(wav, g = 44100, output = "Wave")
      }
      wav@left = round(wav@left)
      mkdirs(dirname(output_filename))
      writeWave(wav, filename = output_filename, extensible = F)
    })
  }
})
```

## Use monitoR to remove silence in sliced clips

```{r}
files = list.files("data/Sound templates", pattern = "*.txt$", full.names=TRUE)
f = files[1]
selection_table = read.table(f, header=TRUE, sep="\t")
freq_lim = c(selection_table$Low.Freq..Hz., selection_table$High.Freq..Hz.)/1000
wave_filename = str_replace(f, ".Table.1.selections.txt", ".wav")
template_id = substr(basename(f), 1, 8)
template_wav = readWave(wave_filename, from = selection_table$Begin.Time..s., to = selection_table$End.Time..s., units="seconds")
template_wav = resamp(template_wav, g = 44100, output = "Wave")
template_duration = length(template_wav@left) / 44100
processed_template_wave_filename = paste0("misc/", basename(wave_filename))
savewav(template_wav, filename = processed_template_wave_filename)
ctemps = makeCorTemplate(processed_template_wave_filename, wl = 300, frq.lim = freq_lim, name = template_id)

for (bird in human_readable_rfids) {
  files_for_bird = Sys.glob(paste0("data/2019 BAR recordings - individual bird clips/", bird, "/*.wav"))
  clips = c()
  if (length(files_for_bird) == 0) {
    next
  }
  for (i in 1:length(files_for_bird)) {
    f = files_for_bird[i]
    cscores = corMatch(f, ctemps)
    cdetects = findPeaks(cscores)
    detects = getDetections(cdetects)
    if (i == 1) {
      plot(cdetects)
    }
    if (nrow(detects) > 1) {
      for (j in 1:nrow(detects)) {
        start_time = detects$time[j]
        clip = readWave(f, from = start_time, to = start_time + template_duration, units = "seconds")@left
        clips = c(clips, clip)
      }
    }
  }
  if (length(clips) > 10) {
    wav = Wave(clips, samp.rate = 44100, bit = 16)
    output_filename = paste0("data/2019 BAR recordings - individual bird clips/", bird, ".wav")
    writeWave(wav, output_filename, extensible = F)
  }
}
```

